An investigation into using scagnostics to find interesting projections of high-dimensional data

### 2018-10-21 scagnostics.Rmd
Example data scaled everywhere.

```{r datasets}
set.seed(1984)
sphere100 <- sphereData(6, 100) %>% rescale() %>% as_tibble()
sphere1000 <- sphereData(6, 1000) %>% rescale() %>% as_tibble()
sphere100S <- scale(sphere100) %>% as_tibble()
sphere1000S <- scale(sphere1000) %>% as_tibble()

pipe100 <- pipeData(6, 100) %>% rescale() %>% as_tibble()
pipe1000 <- pipeData(6,1000) %>% rescale() %>% as_tibble()
pipe100S <- scale(pipe100) %>% as_tibble()
pipe1000S <- scale(pipe1000) %>% as_tibble()

sin100 <- sinData(6, 100) %>% rescale() %>% as_tibble()
sin1000 <- sinData(6, 1000) %>% rescale() %>% as_tibble()
sin100S <- scale(sin100) %>% as_tibble()
sin1000S <- scale(sin1000) %>% as_tibble()
```

@@

```{r indexTable, echo=FALSE}
i <- 1
pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000",
             "Sphere 100 S", "Sphere 1000 S", "Pipe 100 S", "Pipe 1000 S", "Sine 100 S", "Sine 1000 S")
indexOverview <- tibble(convex=numeric(),
               skinny=numeric(),
               stringy=numeric(),
               dcor2d=numeric(),
               splines2d=numeric(),
               MIC=numeric(),
               TIC=numeric(),
               holes=numeric(),
               name=character(),
               vars=character())
dsL <- list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000,
            sphere100S, sphere1000S, pipe100S, pipe1000S, sin100S, sin1000S)
for(ds in dsL){
  for(vars in list(c("V1","V2"),c("V5","V6"),c("V1","V6"))){
    dprj <- as.matrix(select_(ds, vars[1], vars[2]))
    scagRes <- scagnostics(dprj)
    dcorRes <- dcor2d(dprj[,1], dprj[,2])
    splineRes <- splines2d(dprj[,1], dprj[,2])
    mineRes <- mine(dprj[,1], dprj[,2])
    holesRes <- holes(dprj)
    indexOverview <- add_row(indexOverview, convex=scagRes[,"Convex"],
                    skinny=scagRes[,"Skinny"], stringy=scagRes[,"Stringy"],
                    dcor2d=dcorRes, splines2d=splineRes,
                    MIC=mineRes$MIC, TIC=mineRes$TIC, holes=holesRes,
                    name=pLabels[i], vars=paste0(vars[1],"-",vars[2]))
  }
  i <- i+1
}
knitr::kable(indexOverview,  caption = "Summary of index values.", digits=2) 
```

@@

```{r V1V2toV3V4}
if(!file.exists("cache/V1V2toV3V4.rda")){
  m1 <- matrix(c(1,0,0,0,0,0,0,1,0,0,0,0),ncol=2)
  m2 <- matrix(c(0,0,1,0,0,0,0,0,0,1,0,0),ncol=2)
  #this is silly but seems that first two entries are being ignored, so need some fake entries
  m3 <- matrix(c(1,0,0,0,1,0,0,1,0,0,0,1),ncol=2)
  m4 <- matrix(c(1,1,1,0,0,0,0,1,1,0,0,0),ncol=2)
  t1 <- save_history(sin100,tour_path=planned_tour(list(m3,m4,m1,m2)))
  t1full <- as.list(interpolate(t1))
  fullRes <- getProj(sin100, t1full, "sin", 100) %>%
    rbind(getProj(sin1000, t1full, "sin", 1000)) %>%
    rbind(getProj(sphere100, t1full, "sphere", 100)) %>%
    rbind(getProj(sphere1000, t1full, "sphere", 1000)) %>%
    rbind(getProj(pipe100, t1full, "pipe", 100)) %>%
    rbind(getProj(pipe1000, t1full, "pipe", 1000)) %>%
    rbind(getProj(sin100S, t1full, "sin-S", 100)) %>%
    rbind(getProj(sin1000S, t1full, "sin-S", 1000)) %>%
    rbind(getProj(sphere100S, t1full, "sphere-S", 100)) %>%
    rbind(getProj(sphere1000S, t1full, "sphere-S", 1000)) %>%
    rbind(getProj(pipe100S, t1full, "pipe-S", 100)) %>%
    rbind(getProj(pipe1000S, t1full, "pipe-S", 1000))
  save(t1, t1full, fullRes, file = "cache/V1V2toV3V4.rda")
} else {
  load("cache/V1V2toV3V4.rda")
}
```

```{r plotV1V2toV3V4, fig.width=6, fig.height=7, fig.cap="Short tour paths of uninformed views, moving from the projection onto V1-V2 to the projection onto V3-V4, for all considered datasets and index functions. Small samples (p=100) are shown in red, large samples (p=1000) in blue."}
fullResMelt <- melt(fullRes, id=c("t", "name", "size")) %>%
  mutate(sphered = str_detect(name,"-")) %>%
  mutate(name=str_replace(name,"-S",""))
ggplot(fullResMelt, aes(x=t, y=value)) +
  geom_line(aes(color=factor(size), linetype = factor(sphered))) + 
  facet_wrap(variable~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +
  guides(color=FALSE)
```

@@

```{r plannedtourpath}
if(!file.exists("cache/plannedtourpath.rda")){
  m1 <- matrix(c(1,0,0,0,0,0,0,1,0,0,0,0),ncol=2)
  m2 <- matrix(c(0,0,0,0,1,0,0,0,0,0,0,1),ncol=2)
  #this is silly but seems that first two entries are being ignored, so need some fake entries
  m3 <- matrix(c(1,0,0,0,1,0,0,1,0,0,0,1),ncol=2)
  m4 <- matrix(c(1,1,1,0,0,0,0,1,1,0,0,0),ncol=2)
  t2 <- save_history(sin100,tour_path=planned_tour(list(m3,m4,m1,m2)))
  t2full <- as.list(interpolate(t2))
  fullResPlanned <- getProj(sin100, t2full, "sin", 100) %>%
    rbind(getProj(sin1000, t2full, "sin", 1000)) %>%
    rbind(getProj(sphere100, t2full, "sphere", 100)) %>%
    rbind(getProj(sphere1000, t2full, "sphere", 1000)) %>%
    rbind(getProj(pipe100, t2full, "pipe", 100)) %>%
    rbind(getProj(pipe1000, t2full, "pipe", 1000)) %>%
    rbind(getProj(sin100S, t2full, "sin-S", 100)) %>%
    rbind(getProj(sin1000S, t2full, "sin-S", 1000)) %>%
    rbind(getProj(sphere100S, t2full, "sphere-S", 100)) %>%
    rbind(getProj(sphere1000S, t2full, "sphere-S", 1000)) %>%
    rbind(getProj(pipe100S, t2full, "pipe-S", 100)) %>%
    rbind(getProj(pipe1000S, t2full, "pipe-S", 1000))
  save(t2, t2full, fullResPlanned, file = "cache/plannedtourpath.rda")
} else {
  load("cache/plannedtourpath.rda")
}
```

```{r plannedtour, fig.width=6, fig.height=7, fig.cap="As Figure \\ref{fig:plotV1V2toV3V4} but moving from the projection onto V1-V2 to the projection onto V5-V6, i.e. starting from random distribution view, and moving into the special view."}
fullResMeltPlanned <- melt(fullResPlanned, id=c("t", "name", "size")) %>%
  mutate(sphered = str_detect(name,"-")) %>%
  mutate(name=str_replace(name,"-S",""))
ggplot(fullResMeltPlanned, aes(x=t, y=value)) +
  geom_line(aes(color=factor(size),linetype = factor(sphered))) + 
  facet_wrap(variable~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +
  guides(color=FALSE)
```

>>>>

```{r datasets}
set.seed(1984)
sphere100 <- sphereData(6, 100) %>% scale() %>% as_tibble()
sphere1000 <- sphereData(6, 1000) %>% scale() %>% as_tibble()

pipe100 <- pipeData(6, 100) %>% scale() %>% as_tibble()
pipe1000 <- pipeData(6,1000) %>% scale() %>% as_tibble()

sin100 <- sinData(6, 100) %>% scale() %>% as_tibble()
sin1000 <- sinData(6, 1000) %>% scale() %>% as_tibble()
```

@@

```{r indexTable, echo=FALSE}
i <- 1
pLabels <- c("Sphere 100", "Sphere 1000", "Pipe 100", "Pipe 1000", "Sine 100", "Sine 1000")
indexOverview <- tibble(convex=numeric(),
               skinny=numeric(),
               stringy=numeric(),
               dcor2d=numeric(),
               splines2d=numeric(),
               MIC=numeric(),
               TIC=numeric(),
               holes=numeric(),
               name=character(),
               vars=character())
dsL <- list(sphere100, sphere1000, pipe100, pipe1000, sin100, sin1000)
for(ds in dsL){
  for(vars in list(c("V1","V2"),c("V5","V6"),c("V1","V6"))){
    dprj <- as.matrix(select_(ds, vars[1], vars[2]))
    scagRes <- scagnostics(dprj)
    dcorRes <- dcor2d(dprj[,1], dprj[,2])
    splineRes <- splines2d(dprj[,1], dprj[,2])
    mineRes <- mine(dprj[,1], dprj[,2])
    holesRes <- holes(dprj)
    indexOverview <- add_row(indexOverview, convex=scagRes[,"Convex"],
                    skinny=scagRes[,"Skinny"], stringy=scagRes[,"Stringy"],
                    dcor2d=dcorRes, splines2d=splineRes,
                    MIC=mineRes$MIC, TIC=mineRes$TIC, holes=holesRes,
                    name=pLabels[i], vars=paste0(vars[1],"-",vars[2]))
  }
  i <- i+1
}
knitr::kable(indexOverview,  caption = "Summary of index values.", digits=2) 
```

@@

```{r V1V2toV3V4}
if(!file.exists("cache/V1V2toV3V4.rda")){
  m1 <- matrix(c(1,0,0,0,0,0,0,1,0,0,0,0),ncol=2)
  m2 <- matrix(c(0,0,1,0,0,0,0,0,0,1,0,0),ncol=2)
  #this is silly but seems that first two entries are being ignored, so need some fake entries
  m3 <- matrix(c(1,0,0,0,1,0,0,1,0,0,0,1),ncol=2)
  m4 <- matrix(c(1,1,1,0,0,0,0,1,1,0,0,0),ncol=2)
  t1 <- save_history(sin100,tour_path=planned_tour(list(m3,m4,m1,m2)))
  t1full <- as.list(interpolate(t1))
  fullRes <- getProj(sin100, t1full, "sin", 100) %>%
    rbind(getProj(sin1000, t1full, "sin", 1000)) %>%
    rbind(getProj(sphere100, t1full, "sphere", 100)) %>%
    rbind(getProj(sphere1000, t1full, "sphere", 1000)) %>%
    rbind(getProj(pipe100, t1full, "pipe", 100)) %>%
    rbind(getProj(pipe1000, t1full, "pipe", 1000))
  save(t1, t1full, fullRes, file = "cache/V1V2toV3V4.rda")
} else {
  load("cache/V1V2toV3V4.rda")
}
```

```{r plotV1V2toV3V4, fig.width=6, fig.height=7, fig.cap="Short tour paths of uninformed views, moving from the projection onto V1-V2 to the projection onto V3-V4, for all considered datasets and index functions. Small samples (p=100) are shown in red, large samples (p=1000) in blue."}
fullResMelt <- melt(fullRes, id=c("t", "name", "size"))
ggplot(fullResMelt, aes(x=t, y=value)) +
  geom_line(aes(color=factor(size))) + 
  facet_wrap(variable~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +
  guides(color=FALSE)
```

@@

```{r plannedtourpath}
if(!file.exists("cache/plannedtourpath.rda")){
  m1 <- matrix(c(1,0,0,0,0,0,0,1,0,0,0,0),ncol=2)
  m2 <- matrix(c(0,0,0,0,1,0,0,0,0,0,0,1),ncol=2)
  #this is silly but seems that first two entries are being ignored, so need some fake entries
  m3 <- matrix(c(1,0,0,0,1,0,0,1,0,0,0,1),ncol=2)
  m4 <- matrix(c(1,1,1,0,0,0,0,1,1,0,0,0),ncol=2)
  t2 <- save_history(sin100,tour_path=planned_tour(list(m3,m4,m1,m2)))
  t2full <- as.list(interpolate(t2))
  fullResPlanned <- getProj(sin100, t2full, "sin", 100) %>%
    rbind(getProj(sin1000, t2full, "sin", 1000)) %>%
    rbind(getProj(sphere100, t2full, "sphere", 100)) %>%
    rbind(getProj(sphere1000, t2full, "sphere", 1000)) %>%
    rbind(getProj(pipe100, t2full, "pipe", 100)) %>%
    rbind(getProj(pipe1000, t2full, "pipe", 1000))
  save(t2, t2full, fullResPlanned, file = "cache/plannedtourpath.rda")
} else {
  load("cache/plannedtourpath.rda")
}
```

```{r plannedtour, fig.width=6, fig.height=7, fig.cap="As Figure \\ref{fig:plotV1V2toV3V4} but moving from the projection onto V1-V2 to the projection onto V5-V6, i.e. starting from random distribution view, and moving into the special view."}
fullResMeltPlanned <- melt(fullResPlanned, id=c("t", "name", "size"))
ggplot(fullResMeltPlanned, aes(x=t, y=value)) +
  geom_line(aes(color=factor(size))) + 
  facet_wrap(variable~name, ncol=3, scales = "free_y", labeller = label_wrap_gen(multi_line=FALSE)) +
  guides(color=FALSE)
```
